---
title: "Final Programming"
author: "Jordan Dever"
date: "2023-12-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(DescTools); library(MASS)
library(BSDA)
library(car)
library(olsrr)
library(ResourceSelection)
library(tidyverse)
setwd("C:/Users/jorda/OneDrive/Documents/R-Studio/STA6443/HW/")
bw=read.csv("birthweight_final.csv", header=TRUE)
```

```{r}
bw$Weight_Gr = as.factor(bw$Weight_Gr)
bw$Black = as.factor(bw$Black)
bw$Married = as.factor(bw$Married)
bw$Boy = as.factor(bw$Boy)
bw$MomSmoke = as.factor(bw$MomSmoke)
bw$Ed = as.factor(bw$Ed)

```

EXERCISE 1

(1)#############################################################################

```{r}
lm.bw=lm(Weight~ . -Weight_Gr, data=bw)
```

Stepwise
```{r}
bw.stepwise<-ols_step_both_p(lm.bw, pent = 0.01, prem = 0.01, details = FALSE)
bw.stepwise
```

Forward
```{r}
bw.forward<-ols_step_forward_p(lm.bw, penter = 0.01, details = F)
bw.forward
```

Backward
```{r}
bw.backward<-ols_step_backward_p(lm.bw, prem = 0.01, details = F) 
bw.backward 
```
Backward selection kept: Black,MomSmoke,MomWtGain
R-squared
```{r}
bw.best.subset<-ols_step_best_subset(lm.bw) 
bw.best.subset
```
Best model on highest adj R is 'Model 6: Black Married Boy MomSmoke Ed MomWtGain'
where adj r = 0.1314.



Stepwise: MomWtGain, MomSmoke, Black
Forward:  MomWtGain, MomSmoke, Black
Backward:  Black,MomSmoke,MomWtGain (Same as above)
Adj R: Black Married Boy MomSmoke Ed MomWtGain

All of the step based model selection methods had the three same exact variables
of MomWtGain, MomSmoke, Black at a discrimination p-value of 0.01. The only model
that was different was the adjusted R squared model because the highest observed
R squared came from a model that included Married, Boy, and Ed. I think the
step based models all having 3 variables could be due to the fact that we used
a p-value of 0.01 rather than 0.05 while we are only looking for the highest
adjusted R squared and the extra variables could potentially significant at a 
slightly higher p-value cut off.


(2)#############################################################################

```{r}
lm.bw.step = lm(Weight ~ MomWtGain + MomSmoke + Black, data= bw)
inf.id = which(cooks.distance(lm.bw.step)>0.115)
lm.bw.step = lm(Weight ~ MomWtGain + MomSmoke + Black, data= bw[-inf.id,])
par(mfrow=c(2,2))
plot(lm.bw.step, which=c(1:4))
summary(lm.bw.step)
```
After removing outliers, Black1 had a p-value of 0.01133 so I will remove it
from the final model.

```{r}
lm.bw.final = lm(Weight ~ MomWtGain + MomSmoke, data= bw[-inf.id,])
par(mfrow=c(2,2))
plot(lm.bw.final, which=c(1:4))
summary(lm.bw.final)
```
Looking at the diagnostics plots, it looks like the final model mostly follows
a normal distribution. There are a few points that trail off between the -3 to -2
quantiles in the QQ plot. The points look randomly distributed and follow
homoscedascity with the exception of a few points. After removing the outlier 
from the cook's distance, we are left with 399 total observations and only
two variables since Black was 0.01133 after removing the outlier.

(3)

The final model has an Adjusted R-squared of 0.1181 which means that 11.81% of
the variance in weight is explained by our model.

(4)

With a p-value of 5.816e-12, our final model is considered useful because at least
one beta is not zero. We can interpret the predictors as for every one unit 
increase in MomWtGain (p=5.57e-11), then the baby's birth weight is expected to 
increase by 14.1 grams (when MomSmoke is constant). If a mother smokes (p=0.00382), 
then the baby's birth weight is expected to decrease by 222.646 grams.

If MomSmokes
Weight = (3402.041-222.646) + 14.1(MomWtGain Value)

If Mom Doesn't Smoke
Weight = 3402.041 + 14.1(MomWtGain Value)




EXERCISE 2#######################################################################

(1)

```{r}
glm.bw = glm(Weight_Gr ~ . -Weight, data = bw, family = 'binomial')
model.null.AIC = glm(Weight_Gr ~ 1, data=bw, family = binomial)
step.bw.AIC<-step(model.null.AIC, scope = list(upper=glm.bw), 
                  direction="both",test="Chisq", trace = F)
summary(step.bw.AIC)
```

```{r}
model.null.BIC = glm(Weight_Gr ~ 1, data=bw, family = binomial)
step.bw.BIC<-step(model.null.BIC, scope = list(upper=glm.bw),
                  direction="both",test="Chisq", trace = F, k=log(nrow(bw)))
summary(step.bw.BIC)
```

AIC: Boy, MomSmoke, Ed, MomAge, MomWtGain
BIC: MomSmoke, MomAge, MomWtGain

Using a p-value of 0.05:
Both AIC and BIC both have MomSmoke, MomAge, and MomWtGain as contributing factors
toward predicting Weight_Gr while only the AIC selection included Boy and Ed as
other factors but both are insignificant. Both Ed and Boy have a p-value greater 
than 0.05, and MomAge is greater than 0.01 in AIC. In BIC, all p-values are lower
than 0.01 which is the cut off from our previous exercise.

(2)

```{r}
dev.new(width = 1000, height = 1000, unit = "px")
(inf.id=which(cooks.distance(step.bw.BIC)>0.1))
plot(step.bw.BIC, which = 4, id.n = 5) 
```

```{r}
resid.d<-residuals(step.bw.BIC, type = "deviance")
resid.p<-residuals(step.bw.BIC, type = "pearson")
std.res.d<-residuals(step.bw.BIC, type = "deviance")/sqrt(1 - hatvalues(step.bw.BIC)) 
std.res.p <-residuals(step.bw.BIC, type = "pearson")/sqrt(1 - hatvalues(step.bw.BIC)) 

dev.new(width = 1000, height = 1000, unit = "px")
par(mfrow=c(1,2))
plot(std.res.d[step.bw.BIC$model$Weight_Gr==0], col = "red", 
     ylim = c(-3.5,3.5), ylab = "std. deviance residuals", xlab = "ID")
points(std.res.d[step.bw.BIC$model$Weight_Gr==1], col = "blue")

plot(std.res.p[step.bw.BIC$model$Weight_Gr==0], col = "red", 
     ylim = c(-3.5,3.5), ylab = "std. Pearson residuals", xlab = "ID")
points(std.res.p[step.bw.BIC$model$Weight_Gr==1], col = "blue")
```

There were no observations with a Cook's distance larger than the 0.1 cut off, 
so we will use all 400 observations in the model. The distributions look to
have no trends and follow homoskedasticity.

(3)

```{r}
OR.bw = exp(step.bw.BIC$coefficients)
round(OR.bw, 3)
```
The estimate of the odds ratio for MomSmoke1 is 2.377 so the odds for a baby
having low birth weight is 2.377 times the odds for MomSmoke0. The odds ratio
for MomAge is 0.953 so with a one unit increase in MomAge we would expect the odds
of a low birth weight to change by 0.953. The odds ratio for MomWtGain is
0.964 so with a 1 unit increase in MomWtGain we would expect a change in
low birth weight odds by 0.964.

(4)

Based on our final model, an older woman who smokes and gained a lot of weight
during her pregnancy has higher odds in delivering a low birth weight
infant.

(5)


```{r}
bw2=read.csv("birthweight_final.csv", header=TRUE)
bw.fit.prob <- predict(step.bw.BIC, type = "response")
(bw.sample.prop=mean(bw2$Weight_Gr)) # sample proportion 

```
0.4925 is our sample proportion of low birth weight infants.

(6)

```{r}
bw.pred.class <- ifelse(bw.fit.prob > bw.sample.prop, 1, 0) 
cbind(bw[c("MomSmoke", "MomAge", "MomWtGain", "Weight_Gr")], bw.fit.prob, bw.pred.class)
```

```{r}
mean(bw2$Weight_Gr != bw.pred.class)
```
There is a misclassification rate of 0.355 (35.5%) with our sample proportion
cut off of 0.4925.

(7)

```{r}
hoslem.test(step.bw.BIC$y, fitted(step.bw.BIC), g=10) 
```
Our p-value for the Hosmer & Lemeshow GOF test is 0.3252. This means we can 
fail to reject the null and state that our model is adequate and our model 
fits the data well.



EXERCISE 3#######################################################################

Looking at the results for exercises 1 and 2, we can see that a mother smoking
is definitely a risk factor in reducing infant birth weight. In the linear
regression, mother weight gain during pregnancy played a small role in increasing
infant birth weight (~14 for every one unit increase) while in the logistic
regression model it was determined that women who gained during pregnancy had a 
higher odds of lower birth weight. Mother age also played a minor role in
the logistic regression model. I think what this tells us for both models is that
if a mother smokes (-222 lm, and 2.377 OR in glm) the baby much more likely to
be born with a lower birth weight on average. Comparing the two models for mother
weight gain it is possible that more data than 400 observations could be required
to draw effective conclusions, or there is a threshold on ideal mom weight gain.

If we wanted to implement a low-birthweight prevention program, we would suggest
that pregnant women DO NOT smoke for the health of their baby. We would also
suggest that is it likely healthier for the baby if the mother is younger and is
conscious of their weight gain during pregnancy. Its OK for moms to gain weight
it just could possibly become a negative at a certain point.




