---
title: "Assignment 1"
author: "Jordan Dever"
date: "2024-03-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
college <- read_csv("~/R-Studio/Predictive Modeling/ALL CSV FILES - 2nd Edition/College.csv")
```

#2.
Regression, Inference, n = 500 , p = 3
Classification, Prediction, n =20, p = 13
Regression, Prediction, n = 52, p = 3


#5.
A very flexible approach can be disadvantageous when the model overfits to the training data, then leads to a reduced Test MSE. It can be advantageous to not have a strict shape over the data so we can reduce the bias in our model, but our variance will increase. A less flexible approach may be preferred if we want to make inferences on our data and understand the relationship between X and Y. Less flexible approaches are also less complex and easier to describe compared to very flexible models. 

#6.
In terms of statistics, parametric models usually follow the assumptions of the model while non-parametric models violate some assumptions of its parametric counterpart in order to get as close to f without being too wiggly.
Parametric: Parametric models are much more structured and follow defined shapes that aren’t very flexible. Do not require as many observations to create a parametric model. Parametric models also have reduced computation time.
Non-Parametric: Non-parametric models are very flexible and have a better potential to fit the shape of f due to having a wider range of shapes. A large disadvantage of non-parametric measures is that they require much more observations in order to make an accurate estimate of f.

-------------------------------------------------------------------------------------------------------
#8. 
b)
```{r}
View(college)
summary(college)
```

c)
```{r}
pairs(college[ ,3:13])
```

```{r}
college$Private = as.factor(college$Private)
plot(college$Private, college$Outstate)
```


```{r}
Elite=rep("No",nrow(college))
Elite[college$Top10perc >50]="Yes"
Elite=as.factor(Elite)
college=data.frame(college, Elite)

summary(college$Elite)
```

```{r}
plot(college$Elite, college$Outstate)
```

```{r}
par(mfrow = c(2, 2))
hist(college$Apps)
hist(college$Grad.Rate)
hist(college$S.F.Ratio)
hist(college$Expend)
```
```{r}
college_glm = glm(Private ~.-...1, data = college, family = binomial)
summary(college_glm)
```

I ran a logistic regression on all the variables being a function of Private
and I found that F. Undergrad, Outstate, PhD,perc.alumni, and EliteYes are
all significant predictors, at an alpha level of 0.05, for a university being
Private.

```{r}
college_glm = glm(Private ~  `F.Undergrad` + Outstate + PhD 
                  +  `perc.alumni` + Elite, data = college, family = binomial)
summary(college_glm)
```
When running only those significant predictors, Elite no longer becomes significant
in predicting Private.

#9.
a)
```{r}
auto = read_csv("~/R-Studio/Predictive Modeling/ALL CSV FILES - 2nd Edition/Auto.csv")
auto = na.omit(auto)
summary(auto)
```
Everything is numeric EXCEPT horsepower and name according to the summary. I will
change horsepower into numeric because that does not seem correct.
```{r}
auto$horsepower = as.numeric(auto$horsepower)
auto = na.omit(auto)
summary(auto)
```
NA's were introduced when switching horsepower to numeric but have been removed.
Now only 'name' should be the character and everything else numeric.

b)
Range of auto
```{r}
sapply(auto[, 1:7], range)
#mpg cylinders displacement horsepower weight acceleration year
#[1,]  9.0         3           68         46   1613          8.0   70
#[2,] 46.6         8          455        230   5140         24.8   82


```
c)
Mean of auto
```{r}
sapply(auto[, 1:7], mean)
#mpg    cylinders displacement   horsepower       weight acceleration         year 
#23.445918     5.471939   194.411990   104.469388  2977.584184    15.541327    75.979592 
```

Standard Deviation of auto
```{r}
sapply(auto[, 1:7], sd)
#mpg    cylinders displacement   horsepower       weight acceleration         year 
#7.805007     1.705783   104.644004    38.491160   849.402560     2.758864     3.683737 
```

d)
Subsample Range
```{r}
auto2 <- auto[-c(10:85), ]
sapply(auto2[, 1:7], range)
#mpg cylinders displacement horsepower weight acceleration year
#[1,] 11.0         3           68         46   1649          8.5   70
#[2,] 46.6         8          455        230   4997         24.8   82

```
Subsample Mean
```{r}
sapply(auto2[, 1:7], mean)
#mpg    cylinders displacement   horsepower       weight acceleration         year 
#24.404430     5.373418   187.240506   100.721519  2935.971519    15.726899    77.145570 

```
Subsample Standard Deviation
```{r}
sapply(auto2[, 1:7], sd)
#mpg    cylinders displacement   horsepower       weight acceleration         year 
#7.867283     1.654179    99.678367    35.708853   811.300208     2.693721     3.106217 
```
e)
```{r}
pairs(auto[ ,1:8])
```


```{r}
par(mfrow = c(1, 1))
plot(auto$weight, auto$acceleration)

```
Slight correlation that as weight increases, acceleration decreases

```{r}
plot(auto$weight, auto$mpg)

```
Collelation that as weight increases, miles per gallon decreases

f)
Yes, weight seems to be a good predictor on mpg because there is a downward trend in mpg
especially from 3500+ in weight, the mpg goes from ~20 to ~10.
We can also make a model to see if this is true
```{r}
auto_lm = lm(mpg~weight, data = auto)
summary(auto_lm)
```
Weight is a significant predictor based on this linear model.

#10.
a)
```{r}
#install.packages("ISLR2")
library(ISLR2)
boston = ISLR2::Boston
dim(boston)
?ISLR2::Boston
```
506 rows, 13 columns which means 506 suburbs of Boston listed with 13 variables
to help predict housing value in the suburbs. The explaination of each column can
be found in the help function that was called, while each row is a suburb of Boston.

b)
```{r}
pairs(boston)
```

There is a lot going on in the pairwise scatterplots but some plots that look to be correlated at first glance: zn & crim, indus & nox, lstat & medv, rad & tax and possibily more just hard to see with mark I eyeballs

```{r}
par(mfrow = c(2, 2))
plot(boston$tax, boston$crim)
plot(boston$ptratio, boston$crim)
plot(boston$medv, boston$crim)
plot(boston$rm, boston$crim)
```
 It seems that more crime occurs in the higher tax range (~650 to be specific)
 More crime seems to occur when the pupil-teacher ratio is ~20.
 Most crime occurs in the lower median value range of around ~10000
 More crime around 4-6 room houses.
 
c) 
```{r}
boston_lm = lm(crim ~., data = boston)
summary(boston_lm)
```
Using a linear model to find significant predictors of crim (alpha = 0.05), 
significant predictors were zn, dis, rad, and medv. Positive predictors were
zn and rad while dis and medv were negative predictors of crime per capita.
This tells us that as zn and rad increase then crime is likely to increase as
well. If dis and medv were to increase then crime is likely to decrease. Based
on these four variables alone, a suburb that would experience a lot of crime would
have higher zn and rad while having lower dis and medv.

d)
```{r}
par(mfrow=c(1,1))

hist(boston$crim,breaks=50)
```
 
```{r}
range(boston$crim)
```
Most places have low crime, but the value trails all the way up to ~88 per capita 
crime rate by town

```{r}
hist(boston$tax)
range(boston$tax)

```
Most places have a tax range between ~200-400, there there is a cap from 400-600, 
then a big jump in frequency occurs at the 700 tax range

```{r}
hist(boston$ptratio)
range(boston$ptratio)

```
 There is a very high frequency on the ~20 pupil-teacher ratio which could be 
 influencing our prior look at crime vs ptratio in b.
 
 e)
```{r}
dim(subset(Boston, chas == 1))
# 35 bound the Charles River
```
 35 bound the Charles River
 
 f)
 
```{r}
median(Boston$ptratio)
#19.05
```
 
19.05 median pupil-teacher ratio

g)
```{r}
boston[boston$medv == min(boston$medv), ]
summary(boston)
# Compared to our ranges, crime is in the 3rd quartile, indus is in 3rd quartile,
# nox is in the 3rd quartile, rm is below the first quartile, age is max, 
# dis is below 1st quartile, rad is max, tax is in the 3rd quartile, ptratio is in the 3rd quartile
# and lstat is in the 3rd quartile.
```
Compared to our ranges, crime is in the 3rd quartile, indus is in 3rd quartile,
nox is in the 3rd quartile, rm is below the first quartile, age is max, 
dis is below 1st quartile, rad is max, tax is in the 3rd quartile, ptratio is 
in the 3rd quartile and lstat is in the 3rd quartile.

h)
```{r}
dim(subset(boston, rm > 7))
dim(subset(boston, rm > 8))
# 64 houses with 7+ rooms, and 13 with 8+ rooms.
```
64 houses with 7+ rooms, and 13 with 8+ rooms.

```{r}
summary((subset(boston, rm > 8)))
# Crime is very low with 75% of 8 room homes being less than 1 per capita crime rate
# and the median home value is between 21-50, with a mean of 44!
```

Crime is very low with 75% of 8 room homes being less than 1 per capita crime rate
and the median home value is between 21-50, with a mean of 44!