---
title: "Untitled"
format: html
editor: source
---

We begin by installing `tidyverse`.


```{r}
#| eval: false
install.packages('tidyverse')
```
Load the library tidyverse by calling `library()` function

```{r}
library(tidyverse)
```

`mpg` is a dataframe from `ggplot2` package that contains fuel economy data from 1999 and 2008.
```{r}
print(mpg)

```
```{r}
head(mpg)
tail(mpg)
```

## Get the meta data

Dimentions of a dataframe
Rows, Columns
```{r}
dim(mpg)
```

Number of rows
```{r}
nrow(mpg)
```

Number of columns
```{r}
ncol(mpg)
```

Column names

```{r}
colnames(mpg)
```
class of `mpg`

```{r}
class(mpg)
```

```{r}
mpg_df = as.data.frame(mpg)
```
```{r}
class(mpg_df)
```

```{r}
print(mpg_df)
```
## Subsetting a dataframe

```{r}
mpg_df$year
```
Column Selection using `select()` from function `dplyr`

```{r}
select(.data = mpg, year)
```

```{r}
select(mpg, starts_with('c'))
```

```{r}
select(mpg, ends_with('y'))
```


```{r}
select(mpg, contains('e'))
```
## Filtering rows

```{r}
slice(.data = mpg, 10:15)
```

```{r}
filter(.data = mpg, year == 2008)
```

```{r}
filter(.data = mpg, year == 2008, manufacturer == 'audi')
#(year == 2008 & manufacturer == 'audi') to specify the add instead of the comma
# | -> 'or'
```

```{r}
dt = readr::read_csv("Data/svcckww1jskrstnz_csv.zip")
```

